---
title: "The impact of double blind reviewing\\at EvoLang 11: statistics"
output: html_document
---

# Introduction

Note that the analysis here has changed since the original submission, updating three data points with incorrectly assigned genders and adding an extra analysis with deviation coding instead of dummy coding.  The original script and results can be found in the file 'gendercheck_stats3_original.html'.

# Data

This script uses two data files.

AllConferenceData.csv
---------------------

-  Score.Mean:  Mean raw score given by reviewers (scaled between 0 and 1, hierh = better paper)
-  FirstAuthorGender:  Gender of first author
-  conference:  Which conference the paper was submitted to
-  rank:  scaled rank within each conference (higher = better paper)
-  year:  Year number (1-3, for convenience)
-  review:  Type of review (Single / Double blind)
-  po:  Number of conferences the author has previously submitted to


sameAuthorData_3conferences_Differences.csv
-------------------------------------------

-  X: arbitrary author number
-  e10:  best rank for first-authored paper by this author in EvoLang 10
-  e11:  best rank for first-authored paper by this author in EvoLang 11
-  gender: gender of author
-  student: whether the author was a student *in EvoLang 10*
-  maletype:  The strength of the male typing of the topic of the paper (see supplementary materials)
-  diff10_11:  The difference between e10 and e11, for convenience

This data is also available in a different format in "sameAuthorData_3conferences.csv", which also contains data from EvoLang 9.  In the main analysis, it was decided that there were too few authors who applied to all three conferences to take advantage of this.

# Loading data for first analysis

Load libraries.

```{r warning=F, message=F}
# Load data
library(lattice)
library(ggplot2)
library(gplots)
library(xtable)
library(party)

```

```{r}

# read data
allData = read.csv("AllConferenceData.csv",stringsAsFactors = F)
# relabel factor
allData$FirstAuthorGender = factor(allData$FirstAuthorGender,labels=c("Female","Male"))
allData$review = factor(allData$review,levels = c("Single","Double"))
allData$conference = factor(allData$conference,levels = c("EvoLang9","EvoLang10","EvoLang11"))
allData$student = factor(allData$student)
# get rid of unneed columns
allData = allData[,!names(allData) %in% c("AuthorCode",'maletype','X')]

allData$po = allData$po -1

allData$conference.name = c(EvoLang9="E9", EvoLang10="E10", EvoLang11="E11")[allData$conference]
allData$conference.name = factor(allData$conference.name,levels = c("E9","E10","E11"))


```

Look at the distribution of submissions:

```{r}
table(allData$conference, allData$FirstAuthorGender, allData$student)
```

Look at the distributions of raw scores by conference:

```{r}
# Plot data
# Histograms of raw score means by conference
#pdf("Hist_3conf.pdf", height=4, width=12)
par(mfrow=c(1,3))
for(i in levels(allData$conference)){
  hist(allData[allData$conference==i,]$Score.Mean, xlab="Mean paper score", main=i)
}
#dev.off()
par(mfrow=c(1,1))

```

The raw scores are shifted to the right for EvoLang 10 and 11:

```{r}
summary(aov(Score.Mean ~ conference,data=allData))
```

So we'll use normalised rank instead of raw score.

# Review ranks by gender and student status

```{r}
m = aov(rank~(FirstAuthorGender*conference*student), data = allData)
summary(m)

```

```{r echo=F}

res <- as.data.frame(summary(m)[[1]])
res[is.na(res)] = ""

res$sig = ""
res[res$`Pr(>F)` < 0.09,  ]$sig = "."
res[res$`Pr(>F)` < 0.05,  ]$sig = "*"
res[res$`Pr(>F)` < 0.001,  ]$sig = "**"


rownames(res) = gsub(":"," x ",rownames(res))
rownames(res) = gsub("FirstAuthorGender","First author gender",rownames(res))
rownames(res) = gsub("conference","Conference:",rownames(res))
rownames(res) = gsub("student","Student status",rownames(res))

res$`F value` = as.numeric(res$`F value`)
res$`Pr(>F)` = as.numeric(res$`Pr(>F)`)

tab.display = c("s",'d','f','f','f','g','f')
tab.digits  = c(2 , 3 , 3 , 2 , 2 , 2 ,  2)
tab.label = 'tab:regressionTable'
tab.cation = "Results of a linear regression carried out on the data.  Predicted (scaled) rank of paper by conference, gender of first authors and student status, with a baseline taken from female non-students from EvoLang 9."

res.table = xtable(res, display = tab.display, digits=tab.digits, label = tab.label )

x <- capture.output(out <- print(res.table, type='latex'))
try(cat(out, file="../GenderCheck/writeup/latexBits/rankByGenderStudentConference.tex"))

```

## Post hoc t-tests

Whole data:

```{r}
t.test(rank~FirstAuthorGender, data=allData)
```

Compare genders within each conference:

```{r}
t.test.string = function(tx){
  t = signif(tx$statistic,2)
  df = tx$parameter['df']
  p = signif(tx$p.value,3)
  est = signif(diff(tx$estimate),2)
  
  paste("(difference in means = ",est,", t = ",t,", p = ",p,")",sep = "")
}

# EvoLang 9
t.test.string(t.test(rank~FirstAuthorGender, data=allData[allData$conference=='EvoLang9',]))
# EvoLang 10
t.test.string(t.test(rank~FirstAuthorGender, data=allData[allData$conference=='EvoLang10',]))
# EvoLang 11
t.test.string(t.test(rank~FirstAuthorGender, data=allData[allData$conference=='EvoLang11',]))

```

## Number of times submitted

The variable *po* specifies the number of conferences that the author has submitted to, up to the date of the particular submission. (so 1= first submission year to EvoLang, 2 = submitting for the second conference, 3 = submitting for E11, and has also submitted for E10 and E9). 

```{r}
m.first = aov(rank~(FirstAuthorGender*conference*student*po), 
              data = allData)
summary(m.first)

summary(lm(rank~(FirstAuthorGender*conference) + (student*po),data=allData))
```

We see that there is a weak effect of *po* on rank: more experienced submitters do better.  There's also a weak interaction between student status and po: the effect of experience is greater for non-students.

```{r}
plotmeans(rank~paste(po,FirstAuthorGender,sep='/'), data=allData,
          xlab='Experience of submitting to EvoLang / Gender',
          connect = list(1:2,3:4,5:6))

studentYN = c("N",'Y')[1 + as.numeric(allData$student=="Student")]
plotmeans(rank~paste(po,studentYN,sep='/'), data=allData,
          xlab='Experience of submitting to EvoLang / Student?',
          connect = list(1:2,3:4,5:6))
```


## Decision tree

Use classification trees to explore data.  We've changed the minimum criterion from 95\% to 94\% in order to show some of the structure.

```{r fig.width=12}
x = ctree(rank~FirstAuthorGender+conference+student,data=allData, controls=ctree_control(mincriterion = 0.94) )
plot(x)

```

This is broadly in agreement with regressions above.  It suggests that female first authored papers are generally higher, but ranking of male papers declines in EvoLang 11.


# Data on same authors: differences between E10 and E11

```{r}
sameAuthorData3 = read.csv("sameAuthorData_3conferences_Differences.csv", stringsAsFactors = T)
sameAuthorData3$po = sameAuthorData3$po - 2
```

The original results reported in the paper used a linear regression with dummy coding:

```{r}
summary(lm(diff10_11~gender*student, data=sameAuthorData3))
```

However, the data appears more like an interaction with no main effect.  Indeed, when using deviation coding, this is the case:

```{r}
summary(lm(diff10_11~gender*student,
           data=sameAuthorData3, 
           contrasts=list(
             gender=contr.sum(2)/2,
             student=contr.sum(2)/2)
  ))

```

Check effect of po:

```{r}
summary(lm(diff10_11~gender*student + (po*gender),
           data=sameAuthorData3, 
           contrasts=list(
             gender=contr.sum(2)/2,
             student=contr.sum(2)/2)
  ))

```


## Permutation test

```{r}
#permutation test functions
perm = function(dx){
  # balance sample size 
  min.n = min(table(dx$gender))
  diff(tapply(
    dx$diff10_11,
    sample(dx$gender),
    function(X){
      mean(sample(X,min.n))
    }))
}
perm.test = function(dx){
  trueDiff =diff(tapply(
    dx$diff10_11,
    dx$gender,
    mean))
  permDiff = replicate(10000, perm(dx))
  hist(permDiff, xlab='Difference between mean rise in score (E10-E11)\nbetween male and female first authors (permuted data)')
  abline(v=trueDiff,col=2)
  p = sum(permDiff < trueDiff) / length(permDiff)
  z = (trueDiff - mean(permDiff))/sd(permDiff)
  print(paste("Difference between true and permuted data: z = ",round(z,3),", p = ",round(p,3)))
}
```

```{r}
par(mfrow=c(1,1))
set.seed(6789)
# Permutation test for whole data
perm.test(sameAuthorData3)

# Only for non-students
perm.test(sameAuthorData3[sameAuthorData3$student=="Non-Student",])

# Only for students
perm.test(sameAuthorData3[sameAuthorData3$student=="Student",])
```



# Plots

Plot data:  Violin plots of rank by student status in each conference.

```{r}

source("summarySE.r")



p <- ggplot(allData[complete.cases(allData),], aes(factor(conference.name):factor(student), rank, fill=student))

p <- p + geom_violin() + geom_boxplot(width=0.1) +
  theme(text=element_text(size=20), legend.position="none") +
  scale_y_continuous(name="Score ranking")+
  scale_x_discrete(name="")+
  scale_fill_grey(start = 0.55, end=0.8)
p
pdf("Results_Student_3conf.pdf", width = 12, height= 6)
p
dev.off()

```

Rank by gender

```{r}

# Plot data:  Violin plots of rank by gender in E10 and E11

p2 <- ggplot(allData[!is.na(allData$FirstAuthorGender),], aes(factor(conference.name):factor(FirstAuthorGender), rank, fill=FirstAuthorGender))

p2 <- p2 + geom_violin() + geom_boxplot(width=0.1) +
  theme(text=element_text(size=20), legend.position="none") +
  scale_y_continuous(name="Score ranking")+
  scale_x_discrete(name="")+
  scale_fill_grey(start = 0.55, end=0.8)

p2

pdf("Results_Gender_3conf.pdf", width = 12, height= 6)
p2
dev.off()
```

Improvement between E10 and E11 for paired data.

```{r}
p3 <- ggplot(sameAuthorData3, aes(factor(student):factor(gender), diff10_11, fill=gender))

p3 <- p3 + geom_violin() + geom_boxplot(width=0.1, col=1) +
  theme(text=element_text(size=20), legend.position="none") +
  scale_y_continuous(name="Difference in Score ranking E10 - E11")+
  scale_x_discrete(name="")+
  scale_fill_grey(start = 0.55, end=0.8)
p3
pdf("Improvement_3conf.pdf", width = 12, height= 6)
p3
dev.off()

```